---
title: "Generate Artisanal ACE"
author: "`r params$author`"
format: 
  html:
    page-layout: full
editor: visual
toc: true
toc-location: left
toccolor: blue
urlcolor: blue
embed-resources: true 
execute: 
  echo: false
  message: false
  warning: false
params:
  flag: 'FJ'
  year: 2025
  author: "AUTHOR"
---

```{r}
# Load necessary libraries
library(DBI)
library(odbc)
library(tidyverse)
library(flextable)
source("utils.R")

# Define key vars
country_cd <- tolower(params$flag)
report_year = params$year
this_yr_folder = paste0("./data/report_", as.character(report_year), "_", tolower(country_cd), "/") |>
  tolower()
# Define years to use for tables and figures
yrs_long <- (params$year-4):params$year
yr_range <- paste(min(yrs_long),max(yrs_long), sep="-")
'%nin%' = function(x,y) !x %in% y            

```

```{r data-prep}
# Load all raw data - from t2
data_3615 <- read_and_clean(this_yr_folder, country_cd, r_code = 3615) |>
      mutate(across(where(is.character), tolower))

data_3527 <- read_and_clean(this_yr_folder, country_cd, r_code = 3527)|>
      mutate(across(where(is.character), tolower))

data_3614 <- read_and_clean(this_yr_folder, country_cd, r_code = 3614)|>
      mutate(across(where(is.character), tolower))

ika_trips_file = paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")

if (file.exists(ika_trips_file)){
  data_ika_raw <- tryCatch(
    read.csv(paste0(this_yr_folder, "ikasavea/", country_cd, "_trips_landing_site.csv")),
    error = function(e) NULL
  )
  
  if (exists("data_ika_raw")){
    if ("LandingSite" %in% colnames(data_ika_raw) && nrow(data_ika_raw) > 0){
        data_ika_raw <- data_ika_raw|>
          rename(landing_site_name = LandingSite) |>
          group_by(Country, landing_site_name, Year, Month) |>  
          summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop") |>
          filter(Year == report_year)
    }
    
  }
  
  if (is.null(data_ika_raw) || nrow(data_ika_raw) == 0 || is.null(data_ika_raw)) rm(data_ika_raw)
  }

if (!exists("data_ika_raw") && nrow(data_3615) == 0) {
  cat("No Ikasavea or Tails trip data found for this country/year. Document generation stopped.\n")
  knitr::knit_exit()
}

if (exists("data_ika_raw")){
  
   data_ika_trips <- data_ika_raw |>
    mutate(across(where(is.character), tolower)) |>
    janitor::clean_names() |>
    select(landing_site_name, month,
           number_trips)
  
  df_ika_wide <- data_ika_trips %>%
    complete(landing_site_name, month = 1:12, fill = list(number_trips = 0)) %>%
    mutate(month = month.abb[month]) %>%
    pivot_wider(names_from = month, values_from = number_trips) %>%
    select(landing_site_name, all_of(month.abb))
  
  data_ika_catch_kg <- data_ika_raw |>
    mutate(across(where(is.character), tolower)) |>
    janitor::clean_names() |>
    select(landing_site_name, month, skj = skj_kg, yft = yft_kg, bet = bet_kg)#, alb = alb_kg)
  
  data_ika_catch_kg <- data_ika_catch_kg |>
    pivot_longer(cols = c("skj", "bet", "yft"), names_to = "sp_code", values_to = "sp_kg")
  
  }

both_exist <- exists('data_ika_trips') && nrow(data_ika_trips) > 0 && exists('data_3615') && 
              is.data.frame(get('data_3615')) && nrow(data_3615) > 0
```

::: panel-tabset

## Methodology

The steps for preparing the artisanal catch estimates (ACE) for key tuna species are outlined below. The following tabs showing the catch and effort estimates that are used to prepare the final catch estimates.

It should be noted that the Tails and Ikasavea apps are separate applications collecting a sample of the artisanal fishing activity. Some members only use Ikasavea, others use only Tails and some use both. At this point, we do not have comprehensive surveys in place to scale up how much of the fishing activity is recorded using these ereporting platforms in a reliable and representative manner. Therefore, we are basing the ACE estimates presented here on a combination of reporting catch and effort from these two sources (in case data is stored in both sources we will combine their effort and catch values) as well as local expert knowledge of the activities at different landing sites.

In the future, we hope to progress the Frame Survey concept to refine the estimates of total effort that are used to scale up the sampled catch estimates reported in the respective applications.

**1.** We provide a summary of number of trips by month and landing site recorded in Tails (if any)

**2.** We provide a summary of number of trips by month and landing site recorded in Ikasavea (if any)

**3.** If available data is collected using Ikasavea and TAILS we combine (sum) those values to provide final number of trips based on ereporting platforms

**4.** We compare to "rough" estimates of average monthly landings (trips) per month and landing site provided by members

**5.** The tuna catch by landings site and month is calculated (if data is available in Tails and Ikasavea we combine the data, ie: we sum the catches recorded on these two tools)

**6.** We then multiply the monthly CPUE (catch/trip) by the estimated number of trips, and if those data are missing, we use an average CPUE multiplied by the number of trips

**7.** The artisanal ACE is the sum across all months and landing sites

## Fishing Effort

### Data prep: Effort

The tables below present the number of reported **trips** in `r params$year` by month and landing site, as reported in Tails, Ikasavea and in both platforms together (in case you have data reported in these two sources).

If member provided a spreadsheet with the estimated trips per month at each landing site this will also be presented below. Some landing sites may not have any activity, that is OK. The member-provided estimates of effort by landing site and compared with the Tails (or Ikasavea or both sources combined) reported effort to produce a 'best' estimate of effort.

<details>

<summary> Trips: Tails</summary>

```{r}
#| echo: false
#| results: 'asis'
#| label: Monthly-effort
#| tbl-cap: !expr if(nrow(data_3615) > 0) glue::glue("Number of trips per month and landing site reported in Tails") else ""

# Check if data exists
if (!(nrow(data_3615) == 0)) {
  
  names(data_3615) <- gsub(x = names(data_3615), pattern = "_trips$", replacement = "")  
  
  data_3615 <- data_3615 |>
    rename_with(~ month.abb[match(., tolower(month.abb))], 
              .cols = any_of(tolower(month.abb))) |> 
    select(-entity)
  
  flextable(data_3615) |>  
  width(j=1, width=1.2) |> 
  fontsize(size = 10, part = "all")
}else{
  cat("No fishing trips recorded in Tails for the selected year.")
}
```

</details>

<details>

<summary> Trips: Ikasavea</summary>


```{r}
#| echo: false
#| results: 'asis'
#| label: Monthly-effort-ika
#| tbl-cap: !expr if(!exists('data_ika_trips')) glue::glue("Number of trips per month and landing site reported in Ikasavea") else ""
if (!exists('data_ika_trips')){
  cat("No fishing trips recorded in Ikasavea for the selected year.")
}else{
  flextable(df_ika_wide) |>  
  width(j=1, width=1.2) |> 
  fontsize(size = 10, part = "all")
  }
```
</details>

<details>

<summary>Differences in Landing Sites: Ikasavea + Tails</summary>

If you have artisanal data being collected in Tails and in Ikasavea, you will see a summary table below of all landing sites common to T2 and Ikasavea as well the ones unique to each of these sources (if any). Use this section to identify issues in landing site names between these two sources.

```{r}
#| echo: false
#| eval: !expr both_exist
#| results: 'asis'
#| label: diff_t2_ika

if (both_exist) {

  ls_t2  <- tolower(trimws(unique(data_3615$landing_site_name)))
  ls_ika <- tolower(trimws(unique(df_ika_wide$landing_site_name)))
  
  in_both    <- intersect(ls_t2, ls_ika)
  only_t2    <- setdiff(ls_t2, ls_ika)
  only_ika   <- setdiff(ls_ika, ls_t2)
  
  if (length(only_t2) == 0 && length(only_ika) == 0) {
    cat("✅ All landing site names match perfectly between both sources.\n")
  } else {
    
    # Summary table
    summary_df <- data.frame(
      ` ` = c("Matched in both", "Only in T2 (data_3615)", "Only in Ikasavea"),
      `Count` = c(length(in_both), length(only_t2), length(only_ika)),
      check.names = FALSE
    )
    
    # cat("### Landing site name comparison\n\n")
      flextable(summary_df) |>  
         width(j=1, width=1.2) |> 
        fontsize(size = 10, part = "all")

    
    # Table of mismatches
    if (length(only_t2) > 0 || length(only_ika) > 0) {
      max_len <- max(length(only_t2), length(only_ika))
      mismatch_df <- data.frame(
        `Only in T2 (data_3615)` = c(only_t2,  rep("", max_len - length(only_t2))),
        `Only in Ikasavea`       = c(only_ika, rep("", max_len - length(only_ika))),
        `In both` = c(in_both, rep("", max_len - length(in_both))),
        check.names = FALSE
      )
      # cat("\n### Unmatched landing site names\n\n")
      flextable(mismatch_df) |>  
         width(j=1, width=1.2) |> 
        fontsize(size = 10, part = "all")
    }
  }
}

```

</details>

<details>

<summary>Trips: Ikasavea + Tails</summary>

If you have data being collected through these two ereporting tools you will see the final number of trips considering what was reported on both.

```{r}
#| echo: false
#| results: 'asis'
#| label: tails_ika_summary

# ika and tails combined
if (both_exist) {
  prep <- function(df) {
    df %>%
      mutate(landing_site_name = tolower(trimws(landing_site_name))) %>%
      pivot_longer(-landing_site_name, names_to = "month", values_to = "trips") |>
      mutate(month = tolower(gsub("_trips", "", month)))
  }
  
  combined_t2_ika <- bind_rows(prep(data_3615), prep(df_ika_wide)) %>%
    group_by(landing_site_name, month) %>%
    summarise(trips = sum(trips), .groups = "drop") %>%
    mutate(month = factor(month, levels = tolower(month.abb))) %>%
    pivot_wider(names_from = month, values_from = trips) %>%
    arrange(landing_site_name) |>
    select(landing_site_name, tolower(month.abb))
  
  ft <- flextable(combined_t2_ika) |>  
    width(j=1, width=1.2) |> 
    fontsize(size = 10, part = "all")
  
  flextable::htmltools_value(ft)

} else {
  combined_t2_ika <- data.frame()
}
```

</details>

<details>

<summary>Trips: Local Knowledge</summary>

Read in member-estimated **trips** per month at each landing site. Some landing sites may not have any activity, that is OK. The member-provided estimates of effort by landing site and compared with the Tails reported effort to produce a 'best' estimate of effort.

```{r trips-est-mem}
#| echo: false
#| results: 'asis'
#| label: mem-trips
#| tbl-cap: !expr if(nrow(data_3615) > 0) glue::glue("Member reported estimated trips per month per landing site") else ""

local_knowl_trips_file <-  str_c(this_yr_folder,  'additional_files/est_trips.csv')

if (file.exists(local_knowl_trips_file)){
  local_knowl_trips = read.csv(local_knowl_trips_file)
  flextable(local_knowl_trips)  |>  
    fontsize(size = 10, part = "all") |> 
    width(j=1, width=2)
}

```

</details>

### Best estimate of fishing trips

Now estimate monthly trips, based on the **max** *value either reported by the member or reported in the ereporting tools Tails and Ikasavea*. In case the er have data in Tails and Ikasavea the combination of trips in both sources will be used as your best estimate monthly trips (tab above).

```{r prep-ereporting_trips, echo = F}

# check logic to be applied for ereporting: tails, ikasavea or tails + ikasavea?

if (nrow(combined_t2_ika)>0){ # it means tails + ikasavea
  
  ereporting_trips <- combined_t2_ika |>
    mutate(landing_site_name = str_squish(landing_site_name)) |>
    data.frame()
  ereporting_trips_tag <- c("Trips: Ikasavea + Tails")
  
  }else if (!exists("data_ika_trips") && nrow(data_3615) > 0){ # so it means you only have tails
    ereporting_trips <- data_3615 |>
      mutate(landing_site_name = str_squish(landing_site_name)) |>
      data.frame()
    ereporting_trips_tag <- c("Trips: Tails")
      
  }else if(nrow(data_3615) == 0 && nrow(data_ika_trips) > 0){ # it means you only have ikasavea 
    ereporting_trips <- df_ika_wide |>
      data.frame()
    ereporting_trips_tag <- c("Trips: Ikasavea")
  }else{
    ereporting_trips <- data.frame()
    ereporting_trips_tag <- c("")
  }
```

```{r max.trips}
#| echo: false
#| results: 'asis'
#| label: est-trips-monthly
#| tbl-cap: !expr if(nrow(ereporting_trips) > 0) glue::glue("Best estimate of trips per month at each landing site") else ""
 
if (nrow(ereporting_trips) > 0){
  
  if (file.exists(local_knowl_trips_file)){
    
    # check if columns are correct, if not, flag the issue
    if (!all(c("landing_site_name", "est.trips") %in% colnames(local_knowl_trips))) {
      cat("Your estimated trip file has to have the following columns: 'landing_site_name', 'est.trips'")
      best_estimate_trips <- ereporting_trips
        
    }else{
      
      # if all columns are in the df, groom local_knowl_trips and produce best_estimate_trips
      local_knowl_trips <- local_knowl_trips |>
        mutate(across(where(is.character), tolower)) |>
        mutate(landing_site_name = str_squish(landing_site_name))

      best_estimate_trips = ereporting_trips |> 
        full_join(local_knowl_trips |> 
                    select(landing_site_name, est.trips))
      
      best_estimate_trips[is.na(best_estimate_trips)] <- 0
      best_estimate_trips <- data.frame(best_estimate_trips)
      for( i in 2:13){
        best_estimate_trips[ ,i] = ifelse( best_estimate_trips[ ,i] > best_estimate_trips[ ,14],  
                                           best_estimate_trips[ ,i] , best_estimate_trips[ ,14] )
        }

      # Get rid of sites that only have NAs
      best_estimate_trips$num_na_cols <- rowSums(best_estimate_trips[, 2:13] )

      best_estimate_trips <- best_estimate_trips |>  filter(num_na_cols>0) |>
        select(-c(est.trips, num_na_cols))
    }
    }
  else{
    cat(paste0("There is no member-estimated trips available, so the best estimate fishing is the table above: <b>", ereporting_trips_tag,
             "</b>"))    
    best_estimate_trips <- ereporting_trips
  }
  }

flextable(best_estimate_trips)  |>  
  fontsize(size = 8, part = "all")


```

## Catch Estimates

### Data prep: Catch

```{r prep-catch-data}
# ── Month number → abbreviated name (keeps columns in calendar order) ─────────
month_labels <- c("Jan","Feb","Mar","Apr","May","Jun",
                  "Jul","Aug","Sep","Oct","Nov","Dec")

# ── Helper: pivot one subset to wide format ───────────────────────────────────
make_wide <- function(data) {
  data %>%
    # Sum sp_kg for each landing_site × month combination
    group_by(landing_site_name, month) %>%
    summarise(sp_kg = sum(sp_kg, na.rm = TRUE), .groups = "drop") %>%
    # Convert month numbers to abbreviated names
    mutate(month = month_labels[month]) %>%
    # Ensure all 12 months are present before pivoting
    complete(
      landing_site_name,
      month = month_labels,
      fill  = list(sp_kg = 0)
    ) %>%
    # Pivot to wide
    pivot_wider(
      names_from  = month,
      values_from = sp_kg,
      values_fill = 0          # missing combos → 0
    ) %>%
    # Reorder columns so months appear Jan → Dec (only those present in data)
    select(landing_site_name, any_of(month_labels))
}

# helper to combine ika and tails
combine_wide <- function(df1, df2) {
  if (is.null(df1) && is.null(df2)) return(NULL)
  
  month_cols <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
  
  drop_zero_rows <- function(df) {
    if (is.null(df)) return(NULL)
    df |> filter(if_any(all_of(month_cols), ~ . != 0))
  }
  
  df1 <- drop_zero_rows(df1)
  df2 <- drop_zero_rows(df2)
  
  if (is.null(df1) || nrow(df1) == 0) return(df2)
  if (is.null(df2) || nrow(df2) == 0) return(df1)
  
  full_join(df1, df2, by = "landing_site_name", suffix = c("_a", "_b")) |>
    mutate(across(all_of(paste0(month_cols, "_a")), 
                  ~ coalesce(., 0L))) |>
    mutate(across(all_of(paste0(month_cols, "_b")), 
                  ~ coalesce(., 0L))) |>
    mutate(Jan = Jan_a + Jan_b, Feb = Feb_a + Feb_b,
           Mar = Mar_a + Mar_b, Apr = Apr_a + Apr_b,
           May = May_a + May_b, Jun = Jun_a + Jun_b,
           Jul = Jul_a + Jul_b, Aug = Aug_a + Aug_b,
           Sep = Sep_a + Sep_b, Oct = Oct_a + Oct_b,
           Nov = Nov_a + Nov_b, Dec = Dec_a + Dec_b) |>
    select(landing_site_name, all_of(month_cols))
}

# Check if tails data exists - if yes, reshape and create a df per sp
if (nrow(data_3614) > 0) {
  df_by_species <- data_3614 |>
    mutate(landing_site_name = str_squish(landing_site_name)) |>
    group_by(sp_code) %>%
    group_map(~ make_wide(.x), .keep = FALSE) %>%
    setNames(sort(unique(data_3614$sp_code))) 
  
  df_bet <- df_by_species[["bet"]]
  df_skj <- df_by_species[["skj"]]
  df_yft <- df_by_species[["yft"]]
}

# Check if ika data exists - if yeas, reshape and create a df per sp
if(exists('data_ika_catch_kg') && is.data.frame(get('data_ika_catch_kg'))){
  
  df_by_species_ika <- data_ika_catch_kg %>%
    mutate(landing_site_name = str_squish(landing_site_name)) |>

    group_by(sp_code) %>%
    group_map(~ make_wide(.x), .keep = FALSE) %>%
    setNames(sort(unique(data_ika_catch_kg$sp_code))) 
  
  df_bet_ika <- df_by_species_ika[["bet"]]
  df_skj_ika <- df_by_species_ika[["skj"]]
  df_yft_ika <- df_by_species_ika[["yft"]]
  
  
  # If required (if ika and tails exist), combine catch values from both sources
  if (nrow(data_3614) > 0){
    df_bet_total <- combine_wide(df_bet, df_bet_ika)
    df_skj_total <- combine_wide(df_skj, df_skj_ika)
    df_yft_total <- combine_wide(df_yft, df_yft_ika)
  } else{
    # Just use the ika source if t2 data doesn't exist
    df_bet_total <- df_bet_ika
    df_skj_total <- df_skj_ika
    df_yft_total <- df_yft_ika
    
  }
} else {
  # Just use catch from tails if ika data doesn't exist
  df_bet_total <- df_bet
  df_skj_total <- df_skj
  df_yft_total <- df_yft
}
```

```{r prep-unr-data}
# Calculate total effort for the year by landing site using the ereporting trips
tot_trips_ereport = data.frame(landing_site_name = ereporting_trips$landing_site_name,
                    tot_trips = rowSums(ereporting_trips[,-1],na.rm=T)) |>
                    mutate(landing_site_name = str_squish(landing_site_name))
# Reshape trips
best_estimate_trips_long <- best_estimate_trips |>
  janitor::clean_names() |>
  mutate(landing_site_name = str_squish(landing_site_name)) |>
  pivot_longer(cols = 2:13, names_to = "months", values_to = "tot_trips_best")

ereporting_trips_long <- ereporting_trips |>
  janitor::clean_names() |>
  mutate(landing_site_name = str_squish(landing_site_name)) |>
  pivot_longer(cols = 2:13, names_to = "months", values_to = "tot_trips_spc")

# Check if data exists
species = c("skj", "bet", "yft")

unr_list <- list()
raised_list <- list()
for (sp in species){
  df_name <- paste0("df_", sp, "_total")
  if (exists(df_name) && is.data.frame(get(df_name)) && nrow(get(df_name)) > 0) {
    df <- get(df_name)
    # Now calculate monthly average catch per trip
    avg_sp = data.frame(landing_site_name = df$landing_site_name,
                    tot_catch = rowSums(df[,-1],na.rm=T)) |>
                  left_join(tot_trips_ereport) |>
      mutate(cpd = round(tot_catch/tot_trips,0))
    
    unr_sp <- df  |> 
      left_join(avg_sp |> 
                  select(landing_site_name, 
                         cpd))
    
    # Add result to unraised list
    unr_list[[sp]] <- unr_sp # 
    
    # Calculate raised
    unr_sp_long <- unr_sp |>
      janitor::clean_names() |>
      mutate(landing_site_name = str_squish(landing_site_name)) |>
      pivot_longer(cols = 2:13, names_to = "months", values_to = "unr_catch")
  
    raised_df <- unr_sp_long |>
      left_join(ereporting_trips_long, by = c("landing_site_name", "months")) |>
      left_join(best_estimate_trips_long, by = c("landing_site_name", "months")) |>
      mutate(raised = round(case_when(
        is.na(tot_trips_spc) | tot_trips_spc == 0 ~ cpd * tot_trips_best,  # fallback: no sampling data
        .default = (unr_catch / tot_trips_spc) * tot_trips_best            # main: monthly CPUE × best trips
      ), 0))
            # mutate(raised = round(case_when(unr_catch == 0 & tot_trips_spc == 0 ~ cpd * tot_trips_best,
      #                           .default = (unr_catch/tot_trips_spc) * tot_trips_best), 0))

    raised_df_wide <- raised_df |>
      select(landing_site_name, months, raised) |>
      pivot_wider(id_cols = 1, names_from = "months", values_from = "raised")
    
    # Add result to raised list
    raised_list[[sp]] <- raised_df_wide
    
  }
}
```

**Unraised** tuna catches reported in ereporting tools are illustrated below, in kilograms (kgs).

<details>

<summary>Skipjack: SKJ</summary>

```{r}
#| echo: false
#| results: 'asis'
#| label: data_skj
#| tbl-cap: !expr if ("skj" %in% names(unr_list)) glue::glue("Unraised skipjack catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("skj" %in% names(unr_list)) { 
    flextable(unr_list$skj) |>  
    width(j=1, width=1.2) |> 
    fontsize(size = 10, part = "all")
}
```

</details>

<details>

<summary>Yellowfin: YFT</summary>

```{r}
#| echo: false
#| results: 'asis'
#| label: df_yft_total
#| tbl-cap: !expr if ("yft" %in% names(unr_list)) glue::glue("Unraised yellowfin catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("yft" %in% names(unr_list)) { 
    flextable(unr_list$yft) |>  
    width(j=1, width=1.2) |> 
    fontsize(size = 10, part = "all")
}

```

</details>

<details>

<summary>Bigeye: BET</summary>

```{r}
#| echo: false
#| results: 'asis'
#| label: df_bet_total
#| tbl-cap: !expr if ("bet" %in% names(unr_list)) glue::glue("Unraised bigeye catch (kgs) reported in ereporting tools.") else ""

# Check if data exists
if ("bet" %in% names(unr_list)) { 
    flextable(unr_list$bet) |>  
    width(j=1, width=1.2) |> 
    fontsize(size = 10, part = "all")
}

```

</details>

**Raised** catches are produced using the average catch per trip, raised by the best estimate of trips, to produce a representative estimate of total artisanal tuna catches. It should be **noted** that is there are sites with no catches reported, these are removed from the totals, even if effort was indicated by the member. This should be revisited - and an error message is noted in this document when that occurs.

<details>

<summary>Skipjack: SKJ</summary>

Raised catch estimates for skipjack tuna (kg) in `r params$year`:

```{r raise-skj}
# Remove rows where no catch was reported - as no average exists...

if ("skj" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$skj 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_skj = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |> 
    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all")
}
```

</details>

<details>

<summary>Missing skipjack catch with effort</summary>

```{r missing_skj}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_skj') && nrow(missing_unraised_skj) > 0){
  flextable(missing_unraised_skj) |> colformat_num(j = 2:13, digits = 0) |>  fontsize(size = 8, part = "all")
}
```

</details>

<details>

<summary>Yellowfin: YFT</summary>

Raised catch estimates for yellowfin tuna (kg) in `r params$year`:

```{r raise-yft}
# Remove rows where no catch was reported - as no average exists...

if ("yft" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$yft 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_yft = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |> 
    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all")
}
```

</details>

<details>

<summary>Missing yellowfin catch with effort</summary>

```{r missing_yft}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_yft') && nrow(missing_unraised_yft) > 0){
  flextable(missing_unraised_yft) |> colformat_num(j = 2:13, digits = 0) |>  fontsize(size = 8, part = "all")
}
```

</details>

<details>

<summary>Bigeye: BET</summary>

Raised catch estimates for bigeye tuna (kg) in `r params$year`:

```{r raise-bet}
# Remove rows where no catch was reported - as no average exists...

if ("bet" %in% names(raised_list)) { 
  
  df_unraised <- raised_list$bet 
  df_unraised$no_catch <- rowSums(df_unraised[, 2:13],na.rm=T )
  
  missing_unraised_bet = df_unraised |>  filter(no_catch==0)
  
  df_unraised <- df_unraised |>  
    filter(no_catch>0) |>  
    select(-no_catch)
  
  flextable(df_unraised) |> 
    colformat_num(j = 2:13, digits = 0) |>  
    fontsize(size = 8, part = "all")
}
```

</details>

<details>

<summary>Missing bigeye catch with effort</summary>

```{r missing_bet}
# Rows with no catch but some estimated fishing effort
if(exists('missing_unraised_bet') && nrow(missing_unraised_bet) > 0){
  flextable(missing_unraised_bet) |> colformat_num(j = 2:13, digits = 0) |>  fontsize(size = 8, part = "all")
}
```

</details>

## Final Estimates

Here we take the raised catch estimates for the three key tuna species and sum them across all landing sites and months to produce an annual estimate in metric tonnes (mt).

```{r final.ace}
#| label: raised.ace
#| tbl-cap: Raised and unraised annual artisanal catch estimates in mt

ace <- data.frame()

for (sp in species) {
  
  if (sp %in% names(unr_list)) {
    
    unraised <- unr_list[[sp]] |>
      data.frame() |>
      ungroup() |>
      select(-c(1, 14)) |>
      summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) |>
      rowSums()
    
    raised <- raised_list[[sp]] |>
      data.frame() |>
      ungroup() |>
      select(-1) |>
      summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) |>
      rowSums()
    
  } else {
    unraised <- 0
    raised   <- 0
  }
  
  ace <- bind_rows(ace, data.frame(
    sp       = toupper(sp),
    unraised = round(unraised / 1000, 2),
    raised   = round(raised / 1000, 2)
  ))
}

# Produce final output
ace_final = ace |>
  gather(key='Method', value='ACE', -1) |>
  spread(key = sp, value=ACE)

flextable(ace_final) |>  fontsize(size = 12, part = "all")

```

```{r perc.sp}
#| label: ace.perc
#| tbl-cap: Percentage of the artisanal tuna ACE associated with each of the key species

ace_final$tot = rowSums(ace_final[,-1])
ace_final <- ace_final |>  mutate(SKJ = round(SKJ/tot*100,2),
                      YFT = round(YFT/tot*100,2),
                      BET = round(BET/tot*100,2)) |>
  mutate(across(BET:YFT, ~replace_na(.,0))) |>
  select(-tot)

flextable(ace_final) |>
  fontsize(size = 12, part = "all")

```

:::
